---
title: "musichistoRy sentiment analysis"
author: "Drew Walker"
date: "1/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(devtools)
#devtools::install_github('charlie86/spotifyr')
library(spotifyr)
library(tidyverse)
library(genius)
#devtools::install_github("drew-walkerr/musichistoRy")
library(musichistoRy)
library(tictoc)
library(syuzhet)
library(knitr)
```

```{r, loading dataframes}
tic()
myhistory <- get_history("thedrewwalker")
toc() 
```

```{r, history and audiofeatures}
tic()
myhistory_snip <- myhistory %>% 
  head(10) 
toc()

myaudiofeatures <- get_history_audio_features(myhistory_snip,"drew")

```

```{r, get_lyrics}
mysession_df <- get_session_df(myaudiofeatures)

#Should create group_by track dataset that summarizes playcount values and such 
unique_songs <- mysession_df %>% 
  distinct(id, .keep_all = TRUE)

```

```{r, lyrics}
#unique dataset for lyrics 
tic()
mylyrics <- unique_songs %>% 
  add_genius(artist, title, type = "track")
toc()
head(mylyrics, n=100)
kable(mylyrics)
```

# Identifying Language and Translating Lyrics, building list of songs by language

-   <http://code.markedmondson.me/googleLanguageR/>

```{r, Googletranslate}
install.packages("googleLanguageR")
library(googleLanguageR)
gl_auth("googletranslateR-717ce515dfc4.json")
tic()
mylyrics_detected <- mylyrics %>% 
  mutate(language = gl_translate_detect(lyric),
    en_translation = gl_translate(lyric))
toc()

mylyrics_detected2 <- flatten(mylyrics_detected)
save(mylyrics_detected, file = "mylyrics_detected.Rdata")



```

```{r, spanish-deck}

spanish_songs <- mylyrics_detected %>% 
  filter(language$language == "es")

spanish_songs_uniques <- spanish_songs %>% 
  distinct(artist,title)
```

# Creating Anki decks

We'll transform the lyrics data to prepare for the deck import files, and utilize the [reticulate](https://rstudio.github.io/reticulate/index.html) package to interface with [AnkiConnect](https://github.com/FooSoft/anki-connect) to build Anki decks within R.

-   AnkiConnect can build decks, and pull User Statistics, allowing for dashboard anki data as well

```{r, anki-deckbuilding}
library(reticulate)

```

# Using Python to generate decks

-   <https://github.com/kerrickstaley/genanki> genanki may be better, looks like ankiconnect can only make blank decks

-   <https://medium.com/python-in-plain-english/make-flashier-flashcards-automating-anki-with-python-2744ed025366> looks like it could be helpful

```{python, get-anki-connect}
import json
import urllib.request

def request(action, **params):
    return {'action': action, 'params': params, 'version': 6}

def invoke(action, **params):
    requestJson = json.dumps(request(action, **params)).encode('utf-8')
    response = json.load(urllib.request.urlopen(urllib.request.Request('http://localhost:8765', requestJson)))
    if len(response) != 2:
        raise Exception('response has an unexpected number of fields')
    if 'error' not in response:
        raise Exception('response is missing required error field')
    if 'result' not in response:
        raise Exception('response is missing required result field')
    if response['error'] is not None:
        raise Exception(response['error'])
    return response['result']

invoke('createDeck', deck='Song')
result = invoke('deckNames')
print('got list of decks: {}'.format(result))
```

# Using [syuzhet](https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html) package for sentiment analysis

[<https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html>](https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html){.uri}
