---
title: "musichistoRy lyrics analysis"
author: "Drew Walker"
date: "1/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(devtools)
devtools::install_github('charlie86/spotifyr')
library(spotifyr)
library(tidyverse)
library(genius)
devtools::install_github("drew-walkerr/musichistoRy")
library(musichistoRy)
library(tictoc)
library(syuzhet)
library(knitr)
```

```{r, loading dataframes}
tic()
myhistory <- get_history("thedrewwalker")
toc() 
```

```{r, history and audiofeatures}
tic()
myhistory_snip <- myhistory %>% 
  head(10) 
toc()

myaudiofeatures <- get_history_audio_features(myhistory_snip,"drew")

```

```{r, get_lyrics}
mysession_df <- get_session_df(myaudiofeatures)

#Should create group_by track dataset that summarizes playcount values and such 
unique_songs <- mysession_df %>% 
  distinct(id, .keep_all = TRUE)

```

```{r, lyrics}
#unique dataset for lyrics 
tic()
mylyrics <- unique_songs %>% 
  add_genius(artist, title, type = "track")
toc()
head(mylyrics, n=100)
kable(mylyrics)
```

# Identifying Language and Translating Lyrics, building list of songs by language

-   <http://code.markedmondson.me/googleLanguageR/>

```{r, Googletranslate}
install.packages("googleLanguageR")
library(googleLanguageR)
gl_auth("googletranslateR-717ce515dfc4.json")
tic()
mylyrics_detected <- mylyrics %>% 
  mutate(language = gl_translate_detect(lyric),
    en_translation = gl_translate(lyric))
toc()




my_lyrics_translated <- 
  <- gl_translate_detect
```

# Using [syuzhet](https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html) package for sentiment analysis

-   <div>

    </div>

-   <https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html>

```{r, sentiment-analysis}


```

# Creating Anki decks

We'll transform the lyrics data to prepare for the deck import files, and utilize the [reticulate](https://rstudio.github.io/reticulate/index.html) package to interface with [AnkiConnect](https://github.com/FooSoft/anki-connect) to build Anki decks within R.

-   AnkiConnect can build decks, and pull User Statistics, allowing for dashboard anki data as well

```{r, anki-deckbuilding}
library(reticulate)
reticulate::virtualenv_create()

```

```{}
```
